---
title: "Quote ML demo - Big Data - Using RevoScaleR"
output: html_document
---

This is an implementation using RevoScale R functions 
The technical scenario: the data now too large for CRAN R.

First step - stay with local CSV but replace all CRAN R functions with Rx.. functions
See the CRAN implementation for all project background details.


```{r, echo=FALSE, message = FALSE, warning=FALSE}
library(RevoScaleR)
quoteXdfFile <- "xdf/QuoteData.xdf"
quoteXdfFile2 <- "xdf/QuoteData2.xdf"
quoteXdfFile3 <- "xdf/QuoteData3.xdf"

```

# 1. Ingest

```{r}
# read in as a tibble rather than a dataframe
quote <- rxImport(inData = "quotedata.csv", outFile = quoteXdfFile, overwrite = TRUE)
rxGetInfo(quote, getVarInfo = TRUE, numRows=3)
```
# 2. Feature Engineering

```{r}
quote2 <-  rxDataStep(
  inData = quote,
  outFile = quoteXdfFile2,
  overwrite = TRUE,
  varsToDrop = c("TradeKey"),
  reportProgress = 0,
  transforms = list(
    BuySellFlag = ifelse(BuySell == "Buy", 1, 0),
    TradeDate = as.Date(TradeDate),
    splitVar = factor(
      sample(c("Train", "Test"),
      size = .rxNumRows,
      replace = TRUE,
      prob = c(.70, .30)),
      levels = c("Train", "Test"))
    )
)

quote3 <- rxFactors(inData = quote2, 
          outFile = quoteXdfFile3,
          sortLevels = TRUE,
          factorInfo = c("TimeOfDay", "BondCountry", "BondTenor", "CounterpartyCountry", "CounterpartySector", "IsEndOfWeek", "IsEndOfMonth"),
          overwrite = TRUE,
          reportProgress = 0)

rxGetInfo(quote3, getVarInfo = TRUE, numRows=3)
```
# 3. Split into training and test datasets

```{r}
set.seed(1234)
rxSplit(inData = quote3,
           outFilesBase = "xdf/quoteData",
           outFileSuffixes = c("Train", "Test"),
           splitByFactor = "splitVar",
           overwrite = TRUE,
           consoleOutput = FALSE, 
        reportProgress = 0)

train <- RxXdfData("xdf/quoteData.splitVar.Train.xdf")
test <- RxXdfData("xdf/quoteData.splitVar.Test.xdf")

```

# 4a. Model and predict - decision tree model.

```{r}

formula <- "BuySellFlag ~ TimeOfDay + IsEndOfWeek + IsEndOfMonth + 
        BondCountry + BondTenor + 
        CounterpartyCountry +  CounterpartySector"
model_dtree <- rxDTree(formula = formula, data = train, reportProgress = 0)

# let's use the decision tree model to predict
pred_dtree <- rxPredict(model_dtree, test, overwrite = TRUE, reportProgress = 0)

auc_dtree <- rxAuc(rxRoc("BuySellFlag", "BuySellFlag_Pred", test))
paste("AUC of Decision Tree Model is:", sprintf("%.0f %%",auc_dtree * 100))
```

# 4b. Model and predict - logistic regression model.

```{r}
model_logit <- rxLogit(formula = formula, data = train, reportProgress = 0)

pred_logit <- rxPredict(model_logit, 
                        test, 
                        type = "response",
                        predVarNames = "BuySellFlag_Pred_Logit",
                        overwrite = TRUE,
                        reportProgress = 0)

auc_dtree <- rxAuc(rxRoc("BuySellFlag", "BuySellFlag_Pred_Logit", test))
paste("AUC of logistic regression Model is:", sprintf("%.0f %%",auc_dtree * 100))

```

# 5. Compare models

```{r}
rxRocCurve("BuySellFlag",
            predVarNames = c("BuySellFlag_Pred", "BuySellFlag_Pred_Logit"),
            data = test,
            title = "ROC curve - Logistic Regression vs Decision Tree")          
```