---
title: "Quote ML demo - Sparklyr implementation"
output: html_document
---
*Prep: start HDI markspark.  Install sparklyr.  Build Hive table.  Show with Hue*

This is an implementation using sparklyr  
The technical scenario: the data is now on a Azure hadoop HDInsight RServer (Spark) cluster.  This is run from RStudio Server on the edge node.
See the CRAN implementation for all project background details.


```{r, echo=FALSE, message = FALSE, warning=FALSE}
options(repos = "https://mran.microsoft.com/snapshot/2017-05-01") # very necessary
install.packages("sparklyr")
```

# 0. Connect to Spark and set up Spark Context

```{r}
cc <- rxSparkConnect(reset = TRUE, interop = "sparklyr")
sc <- rxGetSparklyrConnection(cc)
```

# 1. Ingest

The data must be in a Hive table for dplyr to operate on it a a Spark  object.

*Show the data in the hive table wuth Hue*

```{r}
quote_data <- RxTextData("/Quote Data.csv", fileSystem = hdfsFS)
quote_data_hive <- RxHiveData(table="QuoteDataTable")
rxDataStep(inData = quote_data, outFile = quote_data_hive, overwrite = TRUE) 

src_tbls(sc) # show all Hive tables
```


```{r}
# create a dplyr tibble pointing to the Spark Hive table
tbl_cache(sc, "QuoteDataTable")
quote_tbl <- tbl(sc, "QuoteDataTable")
```

# 2. Feature Engineering

```{r}
quote_tbl2 <- quote_tbl %>%
  select(-TradeKey) %>% 
  mutate(BuySellFlag = ifelse(BuySell == "Buy", 1, 0),
         TimeOfDay = factor(TimeOfDay),
         BondCountry = factor(BondCountry),
         BondTenor = factor(BondTenor),
         CounterpartyCountry = factor(CounterpartyCountry),
         CounterpartySector = factor(CounterpartySector),
         IsEndOfWeek = factor(IsEndOfWeek),
         IsEndOfMonth = factor(IsEndOfMonth)
         )

```

# 3. Split into training and test datasets

Split data into 70% training and 30% test 

```{r}
quote_tbl3 <- 
  quote_tbl2 %>% 
  sdf_partition(train = 0.7, test = 0.3, seed = 1234)
```

# 4a. Model and predict - linear model.

```{r}
formula <- "BuySellFlag ~ TimeOfDay + IsEndOfWeek + IsEndOfMonth + 
        BondCountry + BondTenor + 
        CounterpartyCountry +  CounterpartySector"

lin_mod <- quote_tbl3$train %>% ml_linear_regression(formula)
summary(lin_mod)

# to do: predict !

```

Close the Spark connection
```{r}
rxSparkDisconnect(cc)
```

